---
title: "ðŸ”® Day 3 - Spell 1: Loading Data Magic"
output: html_document
---

```{r setup-locale, include=FALSE}

```

Welcome young data wizard! Today we learn to open data treasure chests!
Think of CSV files like boxes full of interesting information

---

## Step 1: Load our magical toolbox
These are like apps that give R new powers!

```{r}
# set the locale to English and UTF-8 encoding so that emojis can be displayed correctly
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")

# load the dplyr package
library(dplyr)  # Helps us explore data

print("âœ¨ Toolbox loaded! Ready for data magic!")
```

---

## Step 2: Open our first data treasure chest

ðŸ’¡ **What are we doing?**
Loading data from a CSV file (comma-separated values)
It's like opening a digital treasure chest full of information!

```{r}
# For our class demo, we'll first load the raw creatures data collected from SKY!
creatures_raw <- read.csv("../datasets/creature_of_sky.csv")
print("ðŸŽ‰ Raw creatures data from Sky loaded!")

# ðŸŽˆTry loading the cleaned dataset once we've run the cleaning script:
# creatures_raw <- read.csv("../datasets/creatures.csv")
# uncomment the line above and explore!
```

---

## Step 3: Let's peek inside our treasure chest

Look at the first few rows (like peeking into the box)

```{r}
print("ðŸ‘€ First peek at our raw creature data:")
head(creatures_raw)
```

---

## Step 4: Let's peek inside our treasure chest

Look at the last few rows (like peeking into the box)

```{r}
print("ðŸ‘€ Last peek at our raw creature data:")
tail(creatures_raw)
```

---

## Step 5: Count our treasures

```{r}
# How many responses do we have? (count the rows)
num_rows <- nrow(creatures_raw)
print(paste("ðŸ§™ Total number of creatures:", num_rows))

# How many details about each creature? (count the columns)
num_details <- ncol(creatures_raw)
print(paste("ðŸ“‹ Number of details per creature:", num_details))
```

---

## Step 6: What kind of details do we have?

Show the column names (what information we have about each response)

```{r}
print("ðŸ“ Columns in our raw creatures data:")
names(creatures_raw)
```

## ðŸ’¡ Memory Tip: 
- read.csv() = opens data files
- head() = shows first few rows  
- nrow() = counts rows
- ncol() = counts columns
- names() = shows column names