---
title: "ðŸ”® Day 3 - Spell 5: Histogram Magic"
output: html_document
---

Learn to see the shape of your data!
Histograms show us patterns like mountains and valleys in our data

---

## Step 1: Load tools and data

```{r}
# set the locale to English and UTF-8 encoding so that emojis can be displayed correctly
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")

# load the ggplot2 package
library(ggplot2)  # Our plotting magic toolbox
library(dplyr)
creature_data <- read.csv("../datasets/creatures.csv")

print("ðŸŽ¨ Ready to paint with creature data!")
```

---

## Step 2: Understanding Histograms

ðŸ’¡ **What is a Histogram?**
Histogram is a bar chart that shows the distribution of a variable
It shows how many creatures are in each age range
It shows the shape of the data

```{r}
# Let's look at the ages of our creatures
print("ðŸ” Let's explore creature ages:")
print(creature_data$creature_age)
```

---

## Step 3: Create your first histogram

```{r}
# Show the distribution of creature ages
age_histogram <- ggplot(creature_data, aes(x = creature_age)) +
  geom_histogram(fill = "purple", color = "black") +
  labs(title = "How Old Are Our Magical creatures?",
       x = "Age in Years",
       y = "Number of Creatures")+
  theme_minimal() +
  theme(text = element_text(size = 16),
        plot.title = element_text(size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 19),
        axis.text.x = element_text(angle = 60, hjust = 1))

print(age_histogram)
```

---

## Step 4: Try different data

```{r}
# Look at magic levels
magic_histogram <- ggplot(creature_data, aes(x = magic_power)) +
  geom_histogram(fill = "gold", color = "black") +
  labs(title = "Magic Level Distribution",
       x = "Magic Level",
       y = "Number of Creatures")+
  theme_minimal() +
  theme(text = element_text(size = 16),
        plot.title = element_text(size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 19),
        axis.text.x = element_text(angle = 60, hjust = 1))

print(magic_histogram)
```


---

## ðŸŽˆ Your turn to create!

### Challenge 1: Create a histogram of ages with customized color 

```{r eval=FALSE}
# YOUR CODE HERE:
my_histogram <- ggplot(creature_data, aes(x = ...)) +
  geom_histogram(fill = "...", color = "...") +
  labs(title = "...",
       x = "...",
       y = "...")+
  theme_minimal() +
  theme(text = element_text(size = 16),
        plot.title = element_text(size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 19),
        axis.text.x = element_text(angle = 60, hjust = 1))
```

## ðŸ’¡ Memory Tip:
- ggplot() starts the plot
- aes(x = column) tells what data to use
- geom_histogram() creates the bars
- bins = number of bars to create
- fill = color inside bars, color = outline color
- labs() adds titles and labels