---
title: "🔮 Day 3 - Spell 3: Grouping and Counting Magic"
output: html_document
---

```{r setup-locale, include=FALSE}
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
```

Learn to sort data into groups and count things!
Like sorting candy by color and counting each pile

---

## Step 1: Load data and tools

```{r}
library(dplyr)
creature_data <- read.csv("../datasets/creatures.csv")
print("🎉 Ready to group and count creatures!")
```

---

## Step 2: GROUP BY Magic - Sort into piles

💡 **What is GROUP BY?**
It's like sorting your candy by color - all reds together, all blues together!

```{r}
# Group creatures by their type and count how many of each
creature_counts <- creature_data %>%
  group_by(creature_type) %>%
  summarize(count = n())

print("📊 How many of each creature type:")
print(creature_counts)
```

---

## Step 3: MUTATE Magic - Create new information

💡 **What is MUTATE?**
It's like adding a new sticker to each trading card
You create something new based on what's already there!

```{r}
# Add a column for power category
creatures_with_category <- mutate(creature_data,
                                 power_category = ifelse(magic_power >= 8, "high", "low"))

print("⚡ Creatures with power category:")
print(creatures_with_category)
```

---

## Step 4: More grouping examples

```{r}
# Find the average magic level for each creature type
magic_by_type <- creature_data %>%
  group_by(creature_type) %>%
  summarize(average_magic = mean(magic_power))

print("⚡ Average magic by creature type:")
print(magic_by_type)
```

---

## 🎈 Your turn to practice!

### Challenge 1: Create a new column called "age_group"
If creature_age > 100, it should say "Old", otherwise "Young"

```{r eval=FALSE}
# YOUR CODE HERE:
creatures_with_age_group <- mutate(creature_data, 
                                   age_group = ...)
```

### Challenge 2: Group by age_group and count how many creatures are in each group

```{r eval=FALSE}
# YOUR CODE HERE:
age_counts <- creatures_with_age_group %>%
   group_by(...) %>%
   summarize(...)
```

## 💡 Memory Tip:
- group_by() = sort into groups (like sorting candy)
- summarize() = count or calculate for each group
- mutate() = add new information to each row
- n() = counts how many items in each group
- mean() = calculates the average