---
title: "ðŸ”® Day 3 - Spell 7: Bar Chart Magic"
output: html_document
---

Compare different groups and see who wins!
Bar charts are perfect for competitions and comparisons

---

## Step 1: Load tools and data

```{r}
# set the locale to English and UTF-8 encoding so that emojis can be displayed correctly
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")

# load the ggplot2 package
library(ggplot2)
library(dplyr)

creature_data <- read.csv("../datasets/creatures.csv")

print("ðŸ† Ready for the creature competition!")
```

---

## Step 2: Understanding Bar Charts

ðŸ’¡ **What is a Bar Chart?**
It's like having a competition where each bar shows how many creatures each team has
Longer bars win! Perfect for comparing different groups
Like comparing how many red creatures vs blue creatures vs green creatures you have

```{r}
print("ðŸŽ¯ Let's count our creature types:")
print(table(creature_data$creature_type))
```

---

## Step 3: Create your first bar chart

Count how many of each creature type we have

```{r}
creature_counts <- creature_data %>%
  group_by(creature_type) %>%
  summarize(total = n())

print("ðŸ“Š Creature type counts:")
print(creature_counts)

# Make the bar chart
creature_bar_chart <- ggplot(creature_counts, aes(x = creature_type, y = total)) +
  geom_col(fill = "orange", color = "black") +
  labs(title = "How Many of Each Creature Type?",
       x = "Creature Type",
       y = "Number of Creatures")+
  theme_minimal() +
  theme(text = element_text(size = 16),
        plot.title = element_text(size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 19))

print(creature_bar_chart)
```

---

## Step 4: Compare average magic levels

Which creature type has the highest average magic?

```{r}
magic_by_type <- creature_data %>%
  group_by(creature_type) %>%
  summarize(average_magic = mean(magic_power))

print("âš¡ Average magic by creature type:")
print(magic_by_type)

magic_bar_chart <- ggplot(magic_by_type, aes(x = creature_type, y = average_magic)) +
  geom_col(fill = "purple", color = "black") +
  labs(title = "Which Creature Type is Most Magical?",
       x = "Creature Type",
       y = "Average Magic Level")+
  theme_minimal() +
  theme(text = element_text(size = 16),
        plot.title = element_text(size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 19))

print(magic_bar_chart)
```

---

## ðŸŽˆ Your turn to compete!

### Challenge 1: Create a bar chart showing the average age for each creature type 

```{r eval=FALSE}
# YOUR CODE HERE:
age_totals <- creature_data %>%
  group_by(creature_type) %>%
  summarize(average_age = mean(creature_age))

age_bar_chart <- ggplot(age_totals, aes(x = creature_type, y = average_age)) +
  geom_col(fill = "...", color = "...") +
  labs(title = "...",
       x = "...",
       y = "...")+
  theme_minimal() +
  theme(text = element_text(size = 16),
        plot.title = element_text(size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 19))
```

## ðŸ’¡ Memory Tip:
- geom_col() creates bars for comparisons
- group_by() + summarize() prepares data for bar charts
- fill = color inside bars
- Bars show amounts - longer bars = more of something
- Great for answering "which group has the most?"