---
title: "ðŸ”® Day 2 - Spell 1: Basic Loop Magic"
output: html_document
---

# ðŸŽˆ Activity: Printing Practice
Let's learn how to make R repeat actions automatically!

First, let's see what happens WITHOUT a loop:

```{r}
print("Hello, magical world!")
print("Hello, magical world!")
print("Hello, magical world!")
print("Hello, magical world!")
print("Hello, magical world!")
```

ðŸ˜´ That was tiring to type! Let's use loop magic instead:

## ðŸŒŸ Basic loop - repeat 5 times

```{r}
for (i in 1:5) {
  print("Hello, magical world!")
}
```

---

## ðŸŽ¯ Challenge 1: Print your name 3 times using a for loop
Fill in the blanks, replace the ... with the correct number or string:

```{r}
for (i in 1:...) {
  print("My name is ...")
}
```

Let's see what 'i' actually does:

```{r}
for (i in 1:5) {
  print(paste("This is loop number", i))
}
```

---

## ðŸŽ¯ Challenge 2: Create a countdown from 10 to 1
Hint: Use 10:1 instead of 1:10
Note: paste() is a function that glues strings together

```{r}
for (countdown in ...:...) {
  print(paste("Countdown:", countdown))
}
print("ðŸš€ Blast off!")
```

---

## âœ¨ Advanced Challenge: Print different animals on each loop
Note: c() is a function that creates a collection of items!
TODO: Add a new animal to the list!

```{r}
animals <- c("ðŸ¶ dog", "ðŸ± cat", "ðŸ° rabbit", "ðŸ¸ frog", "ðŸ¦† duck")

for (i in 1:5) {
  print(paste("Animal", i, "is a", animals[i]))
}
```

## ðŸ’¡ Memory Tip
Think of 'i' as a magical counter that changes each time!
The loop runs once for each number in the range you give it.