---
title: "🔮 Day 2 - Spell 9: Visualizing Your Data Magic"
output: html_document
---

# 🎨 Welcome to your magical art studio! Let's turn your dataframe into beautiful pictures!

```{r}
library(ggplot2)
```

---

## 🎯 Challenge 1: Visualize your dataframe from Spell 8
Use the dataframe you created in Spell 8 here!

```{r}
# Example dataframe (replace with yours from Spell 8)
my_data <- data.frame(
  name = c("Pizza", "Ice Cream", "Tacos", "Cookies"),
  rating = c(10, 9, 8, 10),
  category = c("dinner", "dessert", "dinner", "dessert")
)

# Create your first plot with YOUR data!
ggplot(my_data, aes(x = name, y = rating)) +
  geom_col()
```

---

## 🎯 Challenge 2: Add some color magic!
Let's make it more colorful and beautiful

```{r}
ggplot(my_data, aes(x = name, y = rating)) +
  geom_col(fill = "blue")    # Try different colors: "red", "green", "purple", "orange"

# Try this too - different color for each bar!
ggplot(my_data, aes(x = name, y = rating, fill = name)) +
  geom_col()
```

---

## 🎯 Challenge 3: Make it professional with themes!

```{r}
ggplot(my_data, aes(x = name, y = rating)) +
  geom_col(fill = "darkblue") +
  theme_minimal()    # Makes it look clean and modern!

# Try other themes: theme_classic(), theme_dark(), theme_void()
```

---

## 🎯 Challenge 4: Add titles and labels

```{r}
ggplot(my_data, aes(x = name, y = rating)) +
  geom_col(fill = "purple") +
  theme_minimal() +
  labs(
    title = "My Favorite Foods Rating",     # Main title
    x = "Food Items",                       # X-axis label  
    y = "How Much I Like It (1-10)"        # Y-axis label
  )
```

---

## 🎯 Challenge 5: Color by category (if your data has categories)

```{r}
ggplot(my_data, aes(x = name, y = rating, fill = category)) +
  geom_col() +
  theme_minimal() +
  labs(title = "My Food Ratings by Category")
```

---

## ✨ Your Turn: Create your masterpiece!
Take your dataframe from Spell 8 and create the most beautiful 
chart you can imagine! Try different:
- Colors: "red", "blue", "green", "purple", "orange", "pink"
- Themes: theme_minimal(), theme_classic(), theme_dark()
- Titles: Make them fun and descriptive!

```{r eval=FALSE}
# Write your code here:
# my_chart <- ggplot(...
```

## 💡 Pro Tips for this spell:
- aes() tells ggplot what data to use for x, y, and colors
- fill = "color" makes all bars the same color
- fill = column_name makes each category a different color
- theme_minimal() makes charts look professional
- labs() adds titles and labels to make charts easier to understand
- You can save your plot: ggsave("my_chart.png")