---
title: "🔮 Day 2 - Spell 5: Getting Help with the Magic `?` and `??`"
output: html_document
---

---

## 🎯 Challenge 1: Explore the print function with ?

```{r eval=FALSE}
?print
```

Expected: A help page will show up with:
- Description: What print() does
- Usage: How to use it  
- Examples: Code you can try!

---

## 🎯 Challenge 2: Get help with paste

```{r eval=FALSE}
?paste
```

Expected: Help documentation explaining how paste() combines strings

---

## 🎯 Challenge 3: Try the super search with ??
What if you don't know the exact function name? Use ?? to search!

```{r eval=FALSE}
??plotting
```

Expected: A list of ALL functions related to plotting!
This is like asking "Show me everything about plotting"

```{r eval=FALSE}
??statistics
```

Expected: A list of functions related to statistics

---

## 🎯 Challenge 4: Compare ? vs ??
Try both and see the difference!

```{r eval=FALSE}
?plot         # Specific help for plot function
??plot        # All functions related to "plot"
```

💡 Pro Tip: 
- `?` = Get help for ONE specific function (like asking about a specific spell)
- `??` = Search for ALL functions related to a topic (like asking "show me all fire spells")

---

## ✨ Challenge: Pick 3 functions you've learned and explore their help pages
Write down one new thing you learned about each function:

**Function 1:** _______________  
**New thing I learned:** _______________

**Function 2:** _______________  
**New thing I learned:** _______________

**Function 3:** _______________  
**New thing I learned:** _______________