---
title: "Day 4 Extra Challenge Worksheet"
bibliography: ../../references/references.bib
---

# Day 4 Extra Challenge Worksheet

## About the data

In this worksheet, we will be working with the [Pokemon dataset](https://www.kaggle.com/datasets/rounakbanik/pokemon/data), which contains 800+ Pokemon from all seven generations!

<img src=https://upload.wikimedia.org/wikipedia/commons/thumb/9/98/International_Pok%C3%A9mon_logo.svg/2560px-International_Pok%C3%A9mon_logo.svg.png width="500">

The dataset has been reduced with some columns removed. The dataset we will be working with contains the following information: 

- **name:** The English name of the Pokemon
- **pokedex_number:** The entry number of the Pokemon in the National Pokedex
- **percentage_male:** The percentage of the species that are male. Blank if the Pokemon is genderless.
- **type1:** The Primary Type of the Pokemon
- **type2:** The Secondary Type of the Pokemon
- **classification:** The Classification of the Pokemon as described by the Sun and Moon Pokedex
- **capture_rate:** Capture Rate of the Pokemon
- **height_m:** Height of the Pokemon in metres
- **weight_kg:** The Weight of the Pokemon in kilograms
- **hp:** The Base HP of the Pokemon
- **attack:** The Base Attack of the Pokemon
- **defense:** The Base Defense of the Pokemon
- **speed:** The Base Speed of the Pokemon
- **generation:** The numbered generation which the Pokemon was first introduced
- **is_legendary:** Denotes if the Pokemon is legendary.

```{r}
# Load libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, infer)
```

```{r}
# Read in the data
pokemon <- read_csv("data/pokemon.csv")
head(pokemon)
```

## Exercise 1: Creating scatter plots

### 1.1

How does the weight of the Pokemon change as they increase in height? 

Create a scatter plot of the Pokemon in the data set, with `height_m` on the x-axis and `weight_kg` on the y-axis.

What pattern do you observe in the data? Answer in 1-2 sentences.

```{r}
### YOUR CODE HERE
```

> **YOUR ANSWER HERE**

### 1.2

Let's clean up the plot further and make it more "human readable".

Change the x- and y-axis by capitalizing the variable names; put any unit information inside brackets.

In addition, add a title to the plot. The title should **concisely** describe to the audience what the plot is showing (hint: what observation did more make in the previous exercise?)

```{r}
### YOUR CODE HERE
```

### 1.3

Make **at least 1** additional adjustments to the plot. This could be changing the size/colour/shape of the circles inside the scatter plot, changing the font size of the axis labels or title, overall look-and-feel of the plot aesthetics, etc. 

You can reference the below resources (or make use of Google if you want to try something else):

1. Changing plot colour: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
   - Colour Cheat Sheet: https://sape.inf.usi.ch/sites/default/files/ggplot2-colour-names.png
3. Changing the shape/size of the point: http://www.cookbook-r.com/Graphs/Shapes_and_line_types/
4. Changing the overall aesthetic of the plot: https://book.rfortherestofus.com/themes

```{r}
### YOUR CODE HERE
```

### 1.4 (Challenge)

Is the relation between height and weight different for legendary vs non-legendary Pokemon? How could you inspect this visually?

(**Hint:** you can group your dataset by having 1 colour for the legendary Pokemon and 1 colour for the non-legendary ones.)

```{r}
### YOUR CODE HERE
```

## Exercise 2: Understanding Distributions

### 2.1

What is the distribution of **hp points** of all Pokemon in this dataset?

Plot a histogram of this information using a `binwidth` of 10 units. Make sure that the x- and y-axes are human readable and make sure to include a title.

```{r}
### YOUR CODE HERE
```

### 2.2

What is the mean and standard deviation of Pokemons' HP points?

```{r}
### YOUR CODE HERE
```

### 2.3 

Is this distribution symmetric? Are there any outliers in the data? How did you determine if there were or weren't outliers (e.g., did you use a numeric cutoff, inspect the data visually, etc)?

> **YOUR ANSWER HERE**

```{r}
### A CODE CHUNK TO PLAY AROUND WITH THE DATA IF YOU NEED IT
```

## Exercise 3: Sampling

Run the following cell - we will be using this data frame for the remaining questions in this worksheet.

```{r}
set.seed(930) # ensures that the sampling is reproducible
pokemon_encounters <- pokemon |> 
    drop_na(capture_rate, weight_kg) |> 
    uncount(weights = capture_rate) |> 
    slice(sample(1:n())) 

head(pokemon_encounters)
```

### 3.1

Assume the above dataframe represents a **population** of Pokemon in a given region. 

Imagine if you were to encounter and capture a random sample of **30 Pokemon** in this region. Create this sample using code and stored it in a new data frame called `pokemon_sample_30`.

```{r}
set.seed(930) # ensures that the sampling is reproducible

### YOUR CODE HERE

head(pokemon_sample_30)
```

# 3.2

Plot a histogram of the **weight** of this sample of Pokemon. Use a `binwidth` of 15.

Compare the **average weight of this sample of 30 Pokemon** with the **average weight of all Pokemon in this region** by drawing 2 vertical lines on the graph: 
1. A solid red line that represents the average weight of the sample of 30 Pokemon 
2. A dotted blue line that represents the average weight lf all Pokemon in this region

Make sure that your axes are properly labelled and choose an appropriate title for the plot.

```{r}
### YOUR CODE HERE
```

### 3.3

What do you notice about the two means? Is the sample mean higher or lower compared to the population mean?

> **YOUR ANSWER HERE**

### 3.4

Now, if you were take take a bigger sample of **100 Pokemon** from this region. What would the distribution look like? How does the mean of this sample compare to the population mean?

```{r}
set.seed(930) # ensures that the sampling is reproducible

### YOUR CODE HERE
```

> **YOUR ANSWER HERE**

## Acknowledgements

This workshop was originally created by - [**Girls in Data Science**](https://katieburak.github.io/girls-in-DS/) workshop by Dr. Katie Burak, Jenny Lee, and Mona Zhu
